---
title: "Quant II"
subtitle: "Lab 11: Interference"
author: "Junlong Aaron Zhou"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: 
    beamer_presentation:
        theme: "CambridgeUS"
        colortheme: "beaver"
        fig_width: 5
        fig_height: 3 
header-includes: \input{../teX-preamble/preamble}
---


# Outline 
- Interference
    - Interference with known structure
    - General interference
    - Random network
- Contagion
 
# When SUTVA fails
- One of the pioneering works in this field is Manski (1993).
- A linear-in-means model:
$$
Y_i = \alpha + \beta\frac{\sum_{j \in P} Y_j}{n_i} + \gamma X_i + \delta\frac{\sum_{j \in P} X_j}{n_i} + \varepsilon_i.
$$
\pause
- $\beta$: Endogenous effect; $\delta$: Exogenous effect.
- In addition, $\varepsilon_i$ may not be independent to each other due to homophily.
\pause
- Many econometric works are based on this framework (e.g. Bramoullé et al., 2009; Graham, 2014).
- The model can be arbitrarily complicated: dynamics, spatial autoregression, network formation, etc.

# When SUTVA fails
- The outcome-based approach provides a conceptual benchmark.
- When the treatment of others matters: interference.
- When the outcome of others matters: contagion.
- Homophily is often an important source of confounding.
- Challenge: how could we ensure the correctness of the outcome model?
\pause
- We can defend the model via structural approaches.
- Or we can switch to the design-based perspective.

# Interference with known structure
- We assume that the treatment is randomly assigned.
- We possess the knowledge of the "social network" among individuals.
- $i$'s treatment affects $j$'s outcome if and only if there is a tie between them in the network.
\pause
- Partial interference: all units in the same stratum are connected, but no connection between strata.
- Aronow and Samii (2017): with the network we can construct "exposure mapping."
- The network is actually a moderator: $D_i \mapsto W_i$.

# Interference with known structure
- The network allows us to draw the DAG.
- Then everything is identifiable.
- van de Laan (2014) and Ogburn et al. (2018): a general framework to deal with spillover effects in dynamic social networks.
\pause
- The more challenging part is inference.
- We cannot assume that everyone is connected to everyone else.
- Then there is only one observation, $N=1$.
- We have to assume that the spillover is somehow "local."

```{r, echo=FALSE, warning=FALSE, message=FALSE}
library(igraph)
library(igraphdata)
library(interference)
data(USairports)
agraph <- as.undirected(USairports)
amat <- as_adj(agraph, sparse=FALSE)
igraph::write_graph(agraph, './airports_undirected.csv', format='edgelist')
outcomes <- make_dilated_out(amat, make_corr_out, seed=0, hop=1, multipliers=NULL)
tr_vector <- make_tr_vec_permutation(nrow(amat),0.2,R=1,seed=4224)
exposure <- make_exposure_map_AS(amat, tr_vector, hop=1) 
obs_outcome <- rowSums(exposure*t(outcomes))
vars <- data.frame(id=1:755, treatment=tr_vector[1,], y=obs_outcome)
fwrite(vars, './airports_vars_hop1.csv')
# outcomes <- make_dilated_out(amat, make_corr_out, seed=0, hop=2, multipliers=NULL)
# exposure <- make_exposure_map_AS(amat, tr_vector, hop=2) 
# obs_outcome <- rowSums(exposure*t(outcomes))
# vars <- data.frame(id=1:755, treatment=tr_vector[1,], y=obs_outcome)
# fwrite(vars, './airports_vars_hop2.csv')
network <- read_graph('./airports_undirected.csv', format='edgelist', directed=FALSE)
adj_matrix <- as_adj(network, sparse=FALSE)
diag(adj_matrix) <- 0
vars <- read.csv('./airports_vars_hop1.csv')
non_isolates <- which(rowSums(adj_matrix)>0)
adj_matrix <- adj_matrix[non_isolates, non_isolates]
vars <- vars[vars$id %in% non_isolates,]
num_replicates <- 1500
prop_treated <- 0.2
N <- nrow(adj_matrix)
potential_tr_vector <- make_tr_vec_permutation(N, p=prop_treated, R=num_replicates, seed=4224)
exposure <- make_exposure_map_AS(adj_matrix, vars$treatment, hop=1) 
obs_prob_exposure <- make_exposure_prob(potential_tr_vector, adj_matrix, make_exposure_map_AS, list(hop=1))
my_estimates <- estimates(exposure, vars$y, obs_prob_exposure, n_var_permutations = 1000, hop=1) 
```

# Interference with known structure
```{r echo=FALSE, fig.height=7, fig.width=7}
plot(network, vertex.label.cex = 0.4, vertex.color = "red", vertex.frame.color = "red", 
     edge.color = "gray", edge.width = .4, vertex.size = 0.4, edge.arrow.size = .2)
```

# Interference with known structure
```{r, echo=FALSE}
# Horvitz-Thompson estimatorS
my_estimates$tau_ht

# Variance
my_estimates$var_tau_ht
```

# General interference
- But knowing the network is often a very strong assumption.
- Egami (2019): sensitivity analysis for the network's misspecification.
- Recent works proceed under the framework of "general interference."
- There are two questions we are interested in:
    - Is it still possible to estimate the ATE?
    - How can we estimate the interference effect?

# General interference
- For the first question, the answer is yes under some conditions (Aronow, Hudgen and Savje, 2019; Chin, 2017).
- We require the average "perturbation" from changing the treatment status of one unit to be small.
- Notice that the ATE is no longer well-defined, so the limit is actually EATE:
$$
\tau^{*} = \frac{1}{N} \sum_{i=1}^N \tau_i = \frac{1}{N} \sum_{i=1}^N \left[Y_i(1, \mathbf{Z}_{-i}) - Y_i(0, \mathbf{Z}_{-i})\right]
$$
$$
\tau = E_{\mathbf{Z}_{-i}} \tau^{*} = \frac{1}{N} \sum_{i=1}^N \left[E_{\mathbf{Z}_{-i}}Y_i(1, \mathbf{Z}_{-i}) - E_{\mathbf{Z}_{-i}}Y_i(0, \mathbf{Z}_{-i})\right]
$$

# General interference
- For the second question, there is rapid progress in recent years.
- Aronow, Samii and Wang (2020): As the design is known, we can aggregate the individualistic effects in arbitrary ways.
- In spatial experiments, it is natural to do it by distance.
- Just draw circles at each distance $d$ and apply the IPW estimator.
- Inference relies on Stein's method.
 
# General interference
- Wang (2020): the same idea works in TSCS dataset under the assumption of sequential ignorability.
- Researchers have to trade off two sources of confounding: fixed effects vs. interference.
- Papadogeorgou et al. (2020): modeling treatment point process as a stochastic intervention strategy.
- They examine the effect of airstrike in Iraq on rebellions.


# Within-unit interference
- Interference is also a big problem in TSCS data analysis.
- It is not surprising that the treatment in period $t-1$ affects the outcome in period $t$.
- Then the assumption behind the two-way fixed effects model will be violated (no carryover).
- What can we do?
\pause
- If we are fine with abandoning the unit fixed effects: just estimate a MSM.
- Or we have to add more variables into the regression and hope the model is correct.
- A special case is staggered adoption (Strezhnev, 2019; Liu, Wang and Xu, 2020).


# Random network
- All these approaches assume that there is a fixed network and the intervention is at random.
- Another possibility is that the assignment is pre-specified but the network is randomly generated.
- Li et al. (2019)
\begin{center}
\includegraphics[width=.8\linewidth, height=.7\linewidth]{random_network.jpg}
\end{center}

# Random network
- A study using random dorm assignment in Peking University.
- Does living with IMO medalists boost your GPA?
- Notice that this is interference not contagion.
- A design-based approach to understand peer effect in dorms.
\pause
- Two assumption on "local interference:"
    - Partial interference
    - Only attributes matter, not identities
- They propose a Horvitz-Thompson estimator, and prove that it is equivalent to a regression with interaction.

# Random network
\begin{center}
\includegraphics[width=.8\linewidth, height=.7\linewidth]{PKU_result.jpg}
\end{center}
- They also discuss the optimal assignment.
\pause
- Let champions play with champions...

# Network formation
- A relevant question is whether interference affects the network's structure.
- There is a large literature on network formation in social network analysis.
- A familiar approach is to fit a two-way fixed effects model:
$$
Y_{ij} = \mu + \alpha_i + \zeta_j + \beta X_{ij} + \varepsilon_{ij}
$$
- You can further add the time dimension into the model.
- No design-based approaches yet.

# Contagion

- The peril of peer effects (Angrist, 2011)
\pause
- Can you regression your GPA on the average GPA of your dorm?
\pause
- Can you regression your GPA on the average GPA of your roommates?
\pause
- Fowler and Christakis: peer effect of obesity.
\pause
- Egami (2019): a causal framework using static DAG.
- He also has a placebo test on the identification assumption and a debiased estimator.
\pause
- Imai and Jiang (2020) presents it in a meadiatio framework


# Contagion
\begin{center}
\includegraphics[width=.8\linewidth, height=.7\linewidth]{Naoki.jpg}
\end{center}

# Suggested Readings

- For an overview of the interference lecture: [Aronow, Peter; Dean Eckles, Cyrus Samii & Stephanie Zonszein. (2020). “Spillover Effects in Experimental Data.” In James Druckman & Donald Green, Eds. Advances in Experimental Political Science. Cambridge: Cambridge University Press](https://arxiv.org/abs/2001.05444)

- also its reference. 
