---
title: "1. Basic R and Basic Concepts"
author: "Junlong Aaron Zhou [(jlzhou@nyu.edu)](jlzhou@nyu.edu)"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: pdf_document 
fontsize: 12pt
---

# About R
- R is an object-oriented programming language
    - Data
    - Procedures
- Object's procedures can access and modify the data fields of objects.
- If you see a + means a parenthesis or bracket is open.
- R is case sensitive.
- Use / in path names. Not \\.



# Using Third-party Code
- Relevant commands are: `install.packages` and `library`
- Find the appropriate packages and commands with Google and via searching in R:
```
?covariance
??covariance
install.packages("sandwich")
library(sandwich)
library("sandwich")
require(sanwich)
sanwich::vcovHC
?vcovHC
```

# Install from other sources
- If you want to install packages on GitHub:
```
require(devtools)
install_github("wch/ggplot2")
```
- If you have a complied package downloaded on your computer (tar.gz):
- Tools -> Install Packages -> Find the location
- R cmd install *package name*

# Data types
- Character - strings
- Double / Numeric - numbers
- Logical - true/false
- Factor - unordered categorical variables

# Character

 

```{r}
my.name <- "Aaron"
paste("My","name","is","Aaron")
name.sentence <- paste0("My","name","is","Aaron")
as.character(99)
class(my.name)
```



# Numeric

 

```{r}
num <- 99.867
class(num)
round(num, digits=2)
as.numeric("99") + 1
pi
exp(1)
```

# Numeric

- `sin`, `exp`, `log`, `factorial`, `choose`, are some useful mathematical functions 
- You probably noticed that "<-" is an assignment operator
- It lets you store objects and use them later on
- You can also use "=" 
- To remove something, rm(object)
- To remove everything that is stored use rm(list=ls())

# Logical

- The logical type allows us to make statements about truth

 

```{r}
2 == 4
class(2==4)
```

 

```{r}
my.name != num
```

 

```{r}
"34" == 34
```
\footnotesize
- `==`, `!=`, `>`, `<`, `>=`, `<=`, `!`, `&`, `|`, `any`, `all`, etc


# Data Structures

- There are other ways to hold data, though:

    - Vectors/Lists
    - Matrices/Dataframes
    - Array


# Vectors

- Almost everything in R is a vector.

 

```{r}
as.vector(4)
4
```

 

- We can combine elements in vectors with `c`, for concatenate:

 

```{r}
vec <- c("a","b","c")
vec
```

 

```{r}
c(2,3,vec)
```


# More Vectors
- We can index vectors in several ways

 

```{r}
vec[1]
```

 

```{r}
names(vec) <- c("first","second","third")
vec
```

 

```{r}
vec["first"]
```

# Creating Vectors

```{r}
vector1 <- 1:5
vector1

vector1 <- c(1:5,7,11)
vector1

vector2 <- seq(1, 7, 1)
vector2
```
# Creating Vectors

```{r}
cbind(vector1,vector2)
rbind(vector1,vector2)

```

# Missingness

 

```{r}
vec[1] <- NA
vec
```

 

```{r}
is.na(vec)
```

 
```{r}
vec[!is.na(vec)] # vec[complete.cases(vec)]
```

# Lists

- Lists are similar to vectors, but they allow for arbitrary mixing of types and lengths.

 

```{r}
listie <- list(first = vec, second = num)
listie
```

# Lists

```{r}
listie[[1]]
listie$first
```

# Basic Functions

```{r}
a <- c(1,2,3,4,5)
a
sum(a)
max(a)
min(a)
```

# Basic Functions
```{r}
length(a)
length <- length(a)
b <- seq(from=0,to=5,by=.5)
c <- rep(10,27)
d <- runif(100)
```
> More later
# Matrices

- $$A = \begin{pmatrix}1 & 3\\ 2 & 4\end{pmatrix}$$
- $A_{ij}$
- $A_{1,2} = 3$
- $A_{1,\cdot} = (1,3)$

 

```{r}
A <- matrix(c(1,2,3,4),nrow=2,ncol=2)
A
A[1,2]
A[1,]
A[1:2,]
```

# Matrix Operations

- Its very easy to manipulate matrices:

 

```{r}
solve(A) #A^{-1}
```

  

```{r}
10*A
```

# Matrix Operations

```{r}
B<-diag(c(1,2)) #Extract or replace diagonal of a matrix
B
```

 

```{r}
A%*%B
```

# More Matrix Ops.

 

```{r}
t(A) # A'
```

 

```{r}
rbind(A,B)
```

# More Matrix Ops.


```{r}
cbind(A,B)
```

 

```{r}
c(1,2,3)%x%c(1,1) # Kronecker Product
```

- How to generate the OLS estimates with $X$ and $Y$?

# Naming Things

 

```{r}
rownames(A)
```

 

```{r}
rownames(A)<-c("a","b")
colnames(A)<-c("c","d")
A
```

 

```{r}
A[,"d"]
```

# Array

- An array is similar to a matrix in many ways

 

```{r}
array1 <- array(c(1,2,3,4,5,6,7,8), c(2,2,2))
array1
array1[,2,]
```

 

# Dataframes

- The workhorse

- Basically just a matrix that allows mixing of types.

- R has a bunch of datasets 
 
```{r, fig.height=1, fig.width=2}
# data() gives you all the datasets
data(iris)
head(iris)
```

# Dataframes

- But you will generally work with your own datasets
```{r}
getwd()
setwd("/Users/junlong/Dropbox/Teaching/2021_Spring_Quant_2/lab1")
```


-  R can read any number of file types (.csv, .txt, etc.)
```{r}
#.CSV
dat.csv <- read.csv("http://stat511.cwick.co.nz/homeworks/acs_or.csv")
```
# Dataframes
```{r}
#STATA
require(foreign)
dat.data <- read.dta("https://stats.idre.ucla.edu/stat/data/test.dta")
```


# Dataframes
```{r}
# add variables
dat.data[, "intercept"] <- rep(1, nrow(dat.data))
# change the name of a variable
names(dat.data)[6] <- "constant"
# delete variables
dat.data <- dat.data[, -6]
# sort on one variable
dat.data <- dat.data[order(dat.data[, "mpg"]), ]


# remove all missing values
dat.data.complete <- dat.data[complete.cases(dat.data), ]
# Or similarly
dat.dat.nona <- na.omit(dat.data)

dim(dat.data.complete)
dim(dat.dat.nona)

# select a subset
dat.data.subset <- dat.data[dat.data[, "make"] == "AMC", ]
dat.data.subset <- dat.data[1:3, ]
```


# Objects

- Many functions will return objects rather than a single datatype.

 

```{r}
X <- 1:100 
Y <- rnorm(100,X)
out.lm <- lm(Y~X)
class(out.lm)
predict(out.lm)
plot(out.lm)
summary(out.lm)
```

- Objects can have other data embedded inside them

 

```{r}
out.lm$coefficients
```

 

Show results properly using stargazer.
```{r}
library(stargazer)
stargazer(out.lm) #  This code gives you a latex code

stargazer(out.lm, type = "text") # This gives you a table in text
```


You can always includes latex code directly.

\begin{table}[!htbp] \centering
\caption{}
\label{}
\begin{tabular}{@{\extracolsep{5pt}}lc}
\\[-1.8ex]\hline
\hline \\[-1.8ex]
& \multicolumn{1}{c}{\textit{Dependent variable:}} \\
\cline{2-2}
\\[-1.8ex] & Y \\
\hline \\[-1.8ex]
X & 1.004$^{***}$ \\
& (0.003) \\
& \\
Constant & $-$0.172 \\
& (0.200) \\
& \\
\hline \\[-1.8ex]
Observations & 100 \\
R$^{2}$ & 0.999 \\
Adjusted R$^{2}$ & 0.999 \\
Residual Std. Error & 0.991 (df = 98) \\
F Statistic & 85,529.470$^{***}$ (df = 1; 98) \\
\hline
\hline \\[-1.8ex]
\textit{Note:} & \multicolumn{1}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\
\end{tabular}
\end{table}

# Control Flow

- loops
- if/else

# Loops

- for loops - a way to say "do this for each element of the index"
- "this" is defined in what follows the "for" expression

 

```{r}
for(i in 1:5) {
  cat(i*10," ")
}
```

 

```{r}
for(i in 1:length(vec)) { 
  cat(vec[i]," ")
}
```

 

```{r}
for(i in vec) { 
  cat(i," ")
}
```

# If/Else

 

```{r}
if(vec[2]=="b") print("Hello World!")
```

 

```{r}
if(vec[3]=="a") {
  print("Hello World!")
} else {
  print("!dlroW olleH")
}
```


\clearpage{}

# Basic concepts
- When do we want ATT rather than ATE?
- What does conditional independence mean?
- What do we mean by "control?"
- When we have the population, does regression estimate still have uncertainty?
- Lesson One: Whenever you don't understand something, simulate!

# A toy example
```{r dgp}
N_pop <- 1000
X1 <- rnorm(N_pop, 3, 1)
Y0 <- abs(rnorm(N_pop, 5, 2))  
Y1 <- Y0 + rnorm(N_pop, 0, 5) + 4 # PATE = 4 by construction
TE <- Y1 - Y0
ATE <- mean(TE)
```

```{r design uncertainty}
nboots <- 1000
ATE_est <- ATT <- rep(NA, nboots)
for (i in 1:nboots){ 
  D <- rep(0, N_pop)
  D[sample(N_pop, 300)] <- 1
  Y <- D*Y1 + (1-D)*Y0
  data.pop <- data.frame(Y=Y, D=D)
  ATT[i] <- mean(TE[D==1])
  ATE_est[i] <- mean(Y[D==1]) - mean(Y[D==0])
}
plot(density(ATE_est), main = "Bias of the group-mean-difference estimator", xlab = "ATE estimates", xlim = c(0, 10))
abline(v = ATE, lty = 2)
cat("The true ATE is", ATE, ", and the average of ATE estimates is", mean(ATE_est), "\n")
cat("The sampling variance of ATE estimates is", var(ATE_est), "\n")

plot(density(ATT), main = "ATE vs. ATT", xlab = "ATT", xlim = c(0, 10))
abline(v = ATE, lty = 2)
```

```{r FRT}
nboots <- 1000
D <- rep(0, N_pop)
D[sample(N_pop, 500)] <- 1
Y <- D*Y1 + (1-D)*Y0
ATE_original <- mean(Y[D==1]) - mean(Y[D==0])

ATE_boot <- rep(NA, nboots)
for (i in 1:nboots){
  D_boot <- rep(0, N_pop)
  D_boot[sample(N_pop, 500)] <- 1
  ATE_boot[i] <- mean(Y[D_boot==1]) - mean(Y[D_boot==0])
}
plot(density(ATE_boot), main = "Fisher's Randomization Test", xlab = "ATE estimates", xlim = c(-2, 6))
abline(v = ATE_original, lty = 2)
cat("The 95% confidence interval under the null distribution is", quantile(ATE_boot, c(0.025, 0.975)), "\n")
```


```{r total uncertainty, echo=FALSE}
N_sample <- 100
nsamplings <- 4
nboots <- 1000
ATE_s <- ATE_s_var <- ATE_s_mean <- rep(NA, nsamplings)
ATE_est <- ATT <- var_est <- matrix(NA, nsamplings, nboots)

par(mfrow=c(2, 2))
for (j in 1:nsamplings){
  sample_indicator <- sample(N_pop, N_sample)
  Y1_s <- Y1[sample_indicator]
  Y0_s <- Y0[sample_indicator]
  ATE_s[j] <- mean(Y1_s - Y0_s)
  for (i in 1:nboots){ 
    D <- rbinom(N_sample, 1, 0.3) 
    N1_s <- sum(D)
    N0_s <- N_sample - N1_s
    Y <- D*Y1_s + (1-D)*Y0_s
    data.pop <- data.frame(Y=Y, D=D)
    ATT[j, i] <- mean(TE[D==1])
    ATE_est[j, i] <- mean(Y[D==1]) - mean(Y[D==0])
    var_est[j, i] <- var(Y[D==1])/N1_s + var(Y[D==0])/N0_s
  }
  ATE_s_var[j] <- var(ATE_est[j, ])
  ATE_s_mean[j] <- mean(ATE_est[j, ])
  plot(density(ATE_est[j, ]), main = "Density", xlab = "ATE estimates", xlim = c(0, 10))
  abline(v = ATE_s[j], lty = 2)
}


cat("The total variance of ATE estimates is", var(c(ATE_est)), "\n")
cat("The variance of ATE estimates due to design is", mean(ATE_s_var), "\n")
cat("The variance of ATE estimates due to sampling is", var(ATE_s_mean), "\n")
cat("The average variance estimates is", mean(c(var_est)), "\n")

```
