---
title: "Quant II"
subtitle: "Lab 4: Conditioning: Matching, Weighing, and Sensitivity Analysis"
author: "Junlong Aaron Zhou"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: 
    beamer_presentation:
        theme: "CambridgeUS"
        colortheme: "beaver"
header-includes: \input{../teX-preamble/preamble}

---


# Outline
- Blocking and rerandomization
- Matching   
    - Why matching?
    - Various algorithms
    - Asymptotics of matching
- IPW   
    - Why do we love/hate it?
    - Covariate balancing
- What if confounders are unobservable?
    - Placebo
    - Sensitivity
    
# Blocking
- Blocking: covariates adjustment before assignment
- Usually results in more efficient estimates
- Easier to get balance in covariates
- What is the optimal blocking algorithm?

# Rerandomization
- What if your first draw leads to imbalance in covariates?
\pause
- Rubin: draw the assignment again and do not tell anybody!
\pause
- But what is the distribution of the ATE estimates?
- Ding, Li and Rubin (2017): A truncated Gaussian distribution
- Rerandomization can be combined with regression adjustment

# Why matching?
- To approximiate a blocking experiment
- To get rid of model dependence   
Matching is completely nonparametric: $\hat{\tau}_{i} = Y_i - \sum_{\mathcal{M}_i}Y_{i \in \mathcal{M}_i}$.
\pause
- To estimate heterogeneous treatment effects  
Straightforward.
\pause
- To guarantee common support (positivity)  
Suppose we estimate $\tau$ using Lin's approach, then,
$$\hat{\tau} = \bar{Y}_1 - \bar{Y}_0 - (\frac{N_0}{N_0 + N_1} * \hat{\beta_1} + \frac{N_1}{N_0 + N_1} * \hat{\beta_0})'(\bar{X}_1 - \bar{X}_0)$$
(Imbens and Wooldridge, 2009)   
Bias disappears only when $\bar{X}_1 = \bar{X}_0$ (LaLonde, 1986). 
\pause
- Matching cannnot help you get unconfoundedness.

# Basic steps
1. Choose a distance metric
\pause
2. Find matches on your set of covariates/propensity scores, and get rid of non-matches \pause (Warning!)
\pause
3. Check balance in your matched data set
\pause
4. Repeat these steps until your set exhibits acceptable balance
\pause
5. Calulate the ATT/ATE on your matched dataset 

# An example
- Boyd et al. (2010)
- The effect of gender on decision making
- Unit of analysis: the appellate court case
- Treatment: whether there is at least one female in the three judge panel
- Covariates: median ideology, median age, one racial minority, indicator for ideological direction of lower court's decision, indicator for whether a majority of the judges were nominated by Republicans, indicator for whether a majority of the judges on the panel had judicial experience prior to their nomination

```{r data-loading, echo=FALSE}
# Set up data
d <- read.csv("title7_race.csv")
d$LiberalOutcome <- as.integer(d$case_outcome == 'Liberal')
d$Female <- as.integer(d$gender_judge == 'Female')
d$LiberalLowerDirection <- as.integer(d$lower_dir == 'Liberal')
d$Republican <- as.integer(d$party_judge == 'Republican')
d$Experienced <- as.integer(d$jud_experience == 'Experienced')
d$Minority <- as.integer(d$race_judge != 'White')
d$jcs <- d$jcs - min(d$jcs)
d$confirm_yr <- d$confirm_yr - min(d$confirm_yr)
d$Age <- d$dec_year - d$year_birth

case.order <- unique(d$order) # collapse covariates on cases
median.ideo <- unlist(lapply(case.order, function(x) median(d[d$order == x,]$jcs)))
repub.majority <- as.integer(unlist(lapply(case.order, function(x) sum(d[d$order == x,]$Republican) > 1)))
has.minority <- as.integer(unlist(lapply(case.order, function(x) sum(d[d$order == x,]$Minority) > 0)))
maj.experienced <- as.integer(unlist(lapply(case.order, function(x) sum(d[d$order == x,]$Experienced) > 1)))
has.woman <- as.integer(unlist(lapply(case.order, function(x) sum(d[d$order ==x,]$gender_judge == 'Female') > 0)))
liberal.lower.direction <- unlist(lapply(case.order, function(x) unique(d[d$order == x,]$LiberalLowerDirection)))
median.age <- unlist(lapply(case.order, function(x) median(d[d$order == x,]$Age)))
median.confirmation.year <- unlist(lapply(case.order, function(x) median(d[d$order == x,]$confirm_yr)))
ideo.range <- unlist(lapply(case.order, function(x) diff(range(d[d$order == x,]$jcs))))
minorityXmedianIdeo <- median.ideo * has.minority
liberalOutcome <- unlist(lapply(case.order, function(x) unique(d[d$order == x,]$LiberalOutcome)))

# 6 covariates in total
d.new <- data.frame(
  case.order = case.order,
  median.ideo = median.ideo,
  repub.majority = repub.majority,
  has.minority = has.minority,
  maj.experienced = maj.experienced,
  median.age = median.age,
  liberal.lower.direction = liberal.lower.direction,
  has.woman = has.woman,
  liberalOutcome = liberalOutcome)

trt <- d.new$has.woman == 1

means <- apply(d.new[,-8], 2, function(x) tapply(x, trt, mean))
sds <- apply(d.new[-8], 2, function(x) tapply(x, trt, sd))
t.p <- apply(d.new[, -8], 2, function(x) t.test(x[trt], x[!trt])$p.value)
suppressWarnings(ks.p <- apply(d.new[, -8], 2, function(x) ks.test(x[trt], x[!trt])$p.value))
rownames(means) <- rownames(sds) <- c("Treated", "Control")
```

# View Initial Balance
\footnotesize
```{r}
initial.balance <- round(t(rbind(means,t.p,ks.p)),digits=3)[c(2:8),]
initial.balance
```

# Propensity score matching
- Pros: reduce the number of dimensions
- Cons: may not use information in the most efficient way   
\footnotesize
```{r p-score, echo=FALSE, fig.height=4, fig.width=7}
p.model <- glm(trt ~ median.ideo + median.age + 
                 repub.majority + has.minority +
                 maj.experienced + 
                 liberal.lower.direction,
               d.new, family = "binomial")

pscore.logit <-  predict(p.model, type = "response")
hist(pscore.logit)
```

# Propensity score matching
```{r ps-matching, echo=FALSE, fig.height=5, fig.width=7}
d.ctl <- subset(d.new, has.woman == 0)
pscore.logit.ctl <- pscore.logit[!trt]
pscore.logit.trt <- pscore.logit[trt]

d.trt <- subset(d.new, has.woman == 1)
matches <- sapply(pscore.logit.trt, function(x) which.min(abs(pscore.logit.ctl-x)))
d.trt <- rbind(d.trt, d.ctl[matches, ])
pm.logit.mod <- lm(liberalOutcome ~ has.woman + median.ideo + median.age + repub.majority + has.minority +maj.experienced + 
                     liberal.lower.direction, d.trt)
N <- length(matches)
plot(c(pscore.logit.trt, pscore.logit.ctl[matches]), jitter(rep(c(1, 0), c(N,N))), axes = F, ylab= "Treatment and Control", 
     xlab = "Propensity Score")
axis(1)
```

# Nearest Neighbor Matching
- Approximate a blocking experiment
- You can also use `MatchIt` 
\footnotesize
```{r NN-formula, echo=FALSE, message=FALSE}
library(MatchIt)
matching.formula <- as.formula('has.woman ~ 
    median.ideo + median.age + repub.majority +
     has.minority + maj.experienced + liberal.lower.direction')
```

# Nearest Neighbor Matching
\footnotesize
```{r NN-matching, echo=FALSE}
matched.NN <- matchit(matching.formula, method="nearest", data = d.new)
d.NN <- match.data(matched.NN)
t.NN <- apply(d.NN[, c(2:7)], 2, function(x) t.test(x[d.NN$has.woman==1], x[d.NN$has.woman==0])$p.value)
result.NN <- summary(matched.NN)[3][[1]][-1,]
result.NN <- data.frame(cbind(result.NN, t.NN))
result.NN[,c(1:2,8)]
```



# Nearest Neighbor Matching
\footnotesize
```{r NN-plot, message=FALSE, warning=FALSE, fig.height=4}
plot(matched.NN, type="histogram")
```

# Genetic Matching
- Set an objective function and update the distance metric iteratively
$$\sqrt{(X_i - X_j)'(S^{-1/2})'WS^{-1/2}(X_i - X_j)}$$
- Based upon evolutionary algorithm
- It is very slow (especially if you choose a reasonable `pop.size`)
- Can also do it with `Matching` or `GenMatch`

\footnotesize
```{r GE-matching, echo=FALSE, message=FALSE, warning=FALSE}
library(Matching)
library(rgenoud)

gmatch <- GenMatch(d.new$has.woman,
           d.new[,-d.new$has.woman],
           pop.size = 1000,ties=FALSE,print.level=0)
matches <- gmatch$matches[, 2]
match.data <- subset(d.new, trt == 1)
match.data <- rbind(match.data, d.new[matches, ])
trt.factor <- rep(c("Treat","Control"),c(N,N))
means <- apply(match.data[,-1],2,function(x) tapply(x,trt.factor,mean))
sds <- apply(match.data[,-1],2,function(x) tapply(x,trt.factor,sd))
t.p <- apply(match.data[,-1],2,function(x) t.test(x[trt], x[!trt])$p.value)
ks.p <- apply(match.data[,-1],2,function(x) ks.test(x[trt], x[!trt])$p.value)
```

# Genetic Matching
\footnotesize
```{r, GE-result, echo=FALSE, message=FALSE}
#View matches balance
round(t(rbind(means, t.p)), 3)[c(1:6),]
```
- And then you can calculate the effect of interest


# CEM

- CEM creates bins along each covariate dimension (either pre-specified
or automatic)
-  Units lying in the same strata are then matched together
- Curse of dimensionality means that with lots of covariates, we’ll only rarely
have units in the same strata.

# CEM
\footnotesize
```{r CE-matching, message=FALSE}
library(cem)
cem.match <- cem(treatment = "has.woman", 
              data = d.new, 
              drop = "liberalOutcome")
cem.match
```

# CEM

- Hopefully you are lucky and you have more units
- If not, just tweak CEM
\footnotesize
```{r CE-matching1, message=FALSE}

cutpoints <- list( median.ideo=c(0.3,0.5,0.7), 
              median.age= c(60,65))
cem.tweak.match <- cem(treatment = "has.woman", 
                     data = d.new, 
                     drop = "liberalOutcome", cutpoints = cutpoints)
cem.tweak.match
```

# Asymptotics of Matching
- Matching creates extra uncertainty (why?)
- What is the real standard error of $\hat{\tau}$?
\pause
- Roadmap:
    - Abadie and Imbens (2006): asymptotic distribution for NN matching (with replacement)
    - Abadie and Imbens (2011): debiased matching estimator
    - Abadie and Imbens (2008): boostrap doesn't work for matching
    - Abadie and Imbens (2012): matching as a martingale (NN without replacement)
    - Abadie and Imbens (2016): asymptotic distribution for PS matching
    - Otsu and Rai (2017): wild bootstrap for NN matching
    - Bodory et al. (2018): wild bootstrap for PS matching

# Asymptotics of NN Matching
- Denote $E\left[Y_i(D_i) | X_i \right]$ as $\mu_{D_i}(X_i)$, then $Y_i = \mu_{D_i}(X_i) + \epsilon_i$
- Match with K nearest neighbors; replacement is allowed; covariates can be continuous
$$\hat{\tau}_M = \frac{1}{N}\sum_{i=1}^{N}(\widehat{Y}_i(1) - \widehat{Y}_i(0))$$
\pause
- The bias from NN matching can be decomposed into three parts:
$$\hat{\tau}_M - \tau = \overline{\tau(X)} - \tau + E_M + B_M$$
where
$$\overline{\tau(X)} = \frac{1}{N}\sum_{i=1}^{N}(\mu_1(X_i) - \mu_0(X_i))$$
and
$$E_M = \frac{1}{N}\sum_{i=1}^{N}(2D_i - 1)(1+\frac{K_M(i)}{M})\epsilon_i$$

# Asymptotics of NN Matching
- Abadie and Imbens (2006) show that both $\overline{\tau(X)}$ (difference in conditional expectations) and $E_M$ (sum of residuals) are asymptotically unbiased.
- However, the conditional bias relative to  $\overline{\tau(X)}$,
$$B_M = \frac{1}{N}\sum_{i=1}^{N}(2D_i - 1)\left[\frac{1}{M}\sum_{m=1}^{M}(\mu_{1-D_i}(X_i) - \mu_{1-D_i}(X_{j_m(i)}))\right]$$
is not.
\pause
- The bias caused by "mismatch"; it declines very slowly.
- The speed depends on the number of continuous covariates.
- $B_M$ actually converges to an exponential distribution.
- We may estimate $B_M$ directly using the serial estimator proposed by Newey (1995).
- Take-away: do not use bootstrap for NN matching!


# About IPW
- It is actually the Horvitz-Thompson estimator.
\pause
- The duality of IPW and propensity score matching suggests two basic ways of conducting causal inference:
    - Adjust the response surface (matching, regression)
    - Adjust the assignment probability (weighting)
    - Either approach returns unbiased estimate
    - We can combine them to obtain doubly robustness
    
\pause
- IPW can be extended to panel data (dynamic treatment regime).


# The benefits of IPW
- Hirano et al. (2003): the variance of IPW estimators can reach the Cramer-Rao lower bound
- What if we use the real propensity score? 
\pause
- The variance will be larger! (Hahn, 1998)
- Empirical propensity scores take into account all the actual imbalances in the sample


# Caveats for IPW
- It behaves poorly at the "tail" of the support
\pause
- One solution is to stabilize it using the Hajek estimator
\pause  
- Another solution is to drop data at the tail part
- Changes the quantity of interest
\pause
- Ma and Wang (2019): asymptotic distribution for both trimed/untrimed IPW
- They also provide a bias correction method based on resampling

\footnotesize
```{r IPW, echo=FALSE, message=FALSE}
RS.est <- lm(liberalOutcome ~ has.woman + median.ideo + median.age + repub.majority + has.minority + maj.experienced + liberal.lower.direction, d.new)

HA.weight <- trt/pscore.logit + (1 - trt)/(1 - pscore.logit)


IPW.est <- lm(liberalOutcome ~ has.woman, d.new, weights=HA.weight)

DR.est <- lm(liberalOutcome ~ has.woman+median.ideo + median.age +repub.majority+has.minority + 
                      maj.experienced+liberal.lower.direction, d.new, weights=HA.weight)
IPW.means <- t(apply(d.new[, -1], 2, function(x) tapply(x, d.new$has.woman, function(y) weighted.mean(y, weights=HA.weight))))
IPW.means
```



# CBPS and covariates balancing
- IPW balances the covariates, then why don't we direct seek for balance?
- What can you do when the treatment is continuous? 
\pause
- Imai and Ratkovic (2013); Fong, Hazlett and Imai (2018): Covariate Balancing Propensity Score
- Idea: find weights that are orthogonal to $X$, $D$, and their interaction 
$$\sum_{i}^{N} w_i (X^{*}_{i}, D_{i}^{*}, X^{*}_{i}*D_{i}^{*}) = 0 \text{,    } \sum_{i}^{N} w_i = N$$
\pause
- Hainmueller (2012); Hazlett (2015); Arbour and Dimmery (2019)
- Various forms of convex optimization

# CBPS and covariates balancing

```{r balance1, echo=FALSE, message=FALSE}
require(CBPS)
CBPS.score <- CBPS(has.woman ~ median.ideo + median.age + repub.majority + has.minority + maj.experienced + 
                   liberal.lower.direction, d.new)

CBPS.weight <- CBPS.score$weights

CBPS.est <- lm(liberalOutcome ~ has.woman, d.new, weights=CBPS.weight)

CBPS.means <- t(apply(d.new[, -1], 2, function(x) tapply(x, d.new$has.woman, function(y) weighted.mean(y, weights=CBPS.weight))))
CBPS.means
```

```{r balance2, echo=FALSE, message=FALSE}
require(ebal)
eb.score <- ebalance(Treatment = d.new$has.woman, X = cbind(d.new$median.ideo, d.new$median.age, d.new$repub.majority, d.new$has.minority, d.new$maj.experienced, d.new$liberal.lower.direction))

eb.weight <- eb.score$w

eb.means <- t(apply(d.new[, -1], 2, function(x) tapply(x, d.new$has.woman, function(y) weighted.mean(y, weights=eb.weight))))
eb.means
```
 
# Placebo test
- There is no fixed way of conducting placebo test
- Find some variable/observation that should not be affected by the treatment
\pause
- Will the weight of your friends affect yours? How about height?
\pause
- More common in panel data analysis


# Sensitivity analysis
- The basic idea: How influential unobservable confounders have to be to make the estimate insignificant/zero?
\pause
- Remember that confounders must be correlated with both D and Y
- Vary the two correlation coefficients and check how the estimate would change
- Compare the correlation coefficients against observable confounders
- Methods differ in their assumptions on the DGP

# Sensitivity analysis
- First proposed by Rosenbaum and Rubin (1983)
- Imbens (2003): Full parametric model
- Blackwell (2013): Measure selection bias
- Dorie et al. (2016): Semi-parametric test using BART
- Cinelli and Hazlett (2020): Sensitivity from the OVB perspective


# Blackwell (2013)

- Instead of imagining specific uni or multivariate omitted variable, imagine a function which defines the confounding.
- $q(d,x) = E[Y_i(d)|D_i = d, X_i = x] - E[Y_i(d)|D_i = 1 - d, X_i = x]$
- If q is positive units in group $d$ have a higher mean potential outcome under $d$ that those in group $1-d$. 
- So $q$ encodes the selection bias of treatment assignment: it models violations of ignorability.
- After all, confounding means that potential outcomes vary by treatment status. 
- Now, $Y^q_i = Y_i - qPr(1-D_i | X_i)$ and we can redo the analysis.
- Package on CRAN: `causalsens`

# Blackwell (2013)
```{r sen1, echo=FALSE, message=FALSE, fig.height=5}
require(causalsens)
data(lalonde.exp)
ymodel <- lm(re78 ~ treat+age + education + black + hispanic + married + 
               nodegree + re74 + re75 + u74 + u75, data = lalonde.exp)
pmodel <- glm(treat ~ age + education + black + hispanic + married + 
                nodegree + re74 + re75 + u74 + u75, data = lalonde.exp, family = binomial())
alpha <- seq(-4500, 4500, by = 250)
ll.sens <- causalsens(ymodel, pmodel, ~ age + education, data = lalonde.exp, 
                      alpha = alpha, confound = one.sided.att)
plot(ll.sens, type = "raw", bty = "n")
```

# Blackwell (2013)
```{r sen2, echo=FALSE, message=FALSE, fig.height=5}
plot(ll.sens, type = "r.squared", bty = "n")
```

# Cinelli and Hazlett (2020)

- Sensitivity from the omitted variable bias perspective
- Suppose the correct model is $Y = \hat \tau D + \mathbf{X}\hat \beta + \hat \gamma Z + \hat \varepsilon_{full}$
- But $Z$ is unobservable
- So the real model is $Y = \hat \tau_{res} D + \mathbf{X}\hat \beta_{res} + \hat \varepsilon_{res}$
- It is easy to see:
$$
\begin{aligned}
\hat \tau_{res} = & \frac{Cov(D^{\perp \mathbf{X}}, Y^{\perp \mathbf{X}})}{Var(D^{\perp \mathbf{X}})} \\
= & \frac{Cov(D^{\perp \mathbf{X}}, \hat \tau D^{\perp \mathbf{X}} + \hat \gamma Z^{\perp \mathbf{X}})}{Var(D^{\perp \mathbf{X}})} \\
= & \hat \tau + \hat \gamma \frac{Cov(D^{\perp \mathbf{X}}, Z^{\perp \mathbf{X}})}{Var(D^{\perp \mathbf{X}})} \\
= & \hat \tau + \hat \gamma \hat \delta
\end{aligned}
$$

# Cinelli and Hazlett (2020)

- The difference between the correct estimate $\hat \tau$ and the real estimate $\hat \tau_{res}$ has two parts:
    - $\hat \gamma$: the impact of the unobservable
    - $\hat \delta$: the imbalance of the unobservable
- Essentially, the estimate is robust to model misspecification when both $Y$ and $D$ can be largely explained by the observable covariates
- The idea could be extended to the nonlinear case where we use $R^2$ to measure the explanatory power of observable covariates
- Model misspecification is not dependent on the sample size


# Cinelli and Hazlett (2020)
```{r sen3, echo=FALSE, fig.height=3, message=FALSE}
library(sensemakr)
data("darfur")

# runs regression model
model <- lm(peacefactor ~ directlyharmed + age + farmer_dar + herder_dar +
                         pastvoted + hhsize_darfur + female + village, data = darfur)

# runs sensemakr for sensitivity analysis
sensitivity <- sensemakr(model, treatment = "directlyharmed",
                               benchmark_covariates = "female",
                               kd = 1:3)
plot(sensitivity)
```


# Cinelli and Hazlett (2020)
```{r sen4, echo=FALSE, message=FALSE,fig.height=3}
sense_ll <- sensemakr(ymodel, treatment = "treat",
                      benchmark_covariates = "nodegree",
                               kd = 1:3)
plot(sense_ll)
```